---
title: "Fit the phylogenetic regression of RMR by body mass"
author: "Andrew L Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit the phylogenetic regression of RMR by body mass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{rstan}
  %\VignetteDepends{brms}
  %\VignetteDepends{ape}
  \usepackage[utf8]{inputenc}
---


## Setup

```{r}
library(predictFishRMR)
library(rstan)
library(brms)
library(ape)
library(parallel) # run models in parallel

```

## Import and clean data


### Resting metabolic dataset 

Load the raw RMR dataset.  

```{r}

# locate the full directory on this local machine for the MandT data
# in this package
rmr_file_string <- system.file("extdata/MandT.csv",
                                  package = "predictFishRMR")

# load the data
rmr_data <- read.csv( rmr_file_string )
```

The RMR data needs some cleaning so that the species names agree with those in the tree file.

```{r}
rmr_data$spp <- gsub("_", " ", rmr_data$spp) #remove underscore

# some corrections are needed
rmr_data$spp[rmr_data$spp == "Scomberesocidae sp."] <- "Scomberesocidae" # correct this species
rmr_data$spp[rmr_data$spp == "Ambassis sp"] <- "Ambassis sp." # correct this species
rmr_data$spp[rmr_data$spp == "Hypoatherina sp"] <- "Hypoatherina sp." # correct this species
rmr_data$spp[rmr_data$spp == "Monacanthidae sp"] <- "Monacanthidae" # correct this species
rmr_data$spp[rmr_data$spp == "Monacanthidae sp."] <- "Monacanthidae" # correct this species
rmr_data$spp[rmr_data$spp == "Carcharodon carcharius"] <- "Carcharodon carcharias" # correct this species
rmr_data$spp[rmr_data$spp == "Stegostoma tigrinum"] <- "Stegostoma fasciatum"
rmr_data <- subset(rmr_data, spp != "Caulophrynidae sp.") # this family is not included in the phylo tree
rmr_data <- rmr_data[order(rmr_data$spp),]
```

### Species names dataset

Load the species names data.

```{r}
# locate the full directory on this local machine for the species names data
# in this package
sp_file_string <- system.file("extdata/sp_list.csv",
                                  package = "predictFishRMR")

# load the file
sp_names <- read.csv(sp_file_string, sep = ";")
```

Process the species names data. 

```{r}
sp_names$spp <- gsub("_", " ", sp_names$spp) #remove underscore
sp_names$original.name <- gsub("_", " ", sp_names$original.name) #remove underscore
sp_names$sp_all <- gsub("_", " ", sp_names$sp_all) #remove underscore
sp_names$original.name[sp_names$original.name == "Carcharodon carcharius"]<- "Carcharodon carcharias" # correct this species
sp_names$original.name[sp_names$original.name == "Stegostoma tigrinum"]<- "Stegostoma fasciatum" # correct this species
```

### Phylogenetic tree dataset

```{r}
# locate the full directory on this local machine for the tree data
# in this package
tree_file_string <- system.file("extdata/shark_fish.trees",
                                  package = "predictFishRMR")

# load the file
tree <- read.tree(tree_file_string)
```

Process the tree tips for matching. 

```{r}
tree$tip.label <- gsub("_", " ", tree$tip.label) #remove underscore
tree$tip.label <- gsub("-", ",", tree$tip.label) #remove underscore
tree$tip.label[tree$tip.label == "Pseudocaranx dentex ,formaly Longirostrum delicatissimus"] <- "Pseudocaranx dentex (formaly Longirostrum delicatissimus) "
tree$tip.label[tree$tip.label == "Sardinops sagax ,formaly S. caerulea"] <- "Sardinops sagax (formaly S. caerulea)"
```

### Other data processing

AJ - need to rename this section and check how best to present it with Nacho.

```{r}
# Info for Stegostoma_tigrinum comes from Stegostoma_fasciatum
sp_tree <- unique(tree$tip.label)
sp <- unique(rmr_data$spp)

# check point: diff must be 0
setdiff(rmr_data$spp, tree$tip.label)
setdiff(rmr_data$spp, sp_names$original.name)

# --- housekeeping
rmr_data$method <- factor(rmr_data$method)
rmr_data$ram <- factor(rmr_data$ram)
rmr_data$therm <- factor(rmr_data$therm)
rmr_data <- rmr_data[complete.cases(rmr_data),]
```

## Bayesian Linear Mixed / Phylogenetic Model

```{r, eval = FALSE}
# construct covariance matrix of species (Hadfield & Nakagawa, 2010).
A <- ape::vcv.phylo(compute.brlen(tree,method="Grafen"))

# adding species as a factor into our dataset
rmr_data$spp_name<-factor(rmr_data$spp)

# Many species include serveral masurements. Thus, we use a bayesian models.
# General STAN specifications 
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

		
		
		# MODEL
		# We use a bayesian approach to study the relationship
		# across species (phylogeny) while taking into considration
		# intraspaecific variation and multiple observations for
		# several species.



# set model priors
priors =c(
  prior(normal(0,10), "b"),
  prior(normal(0,50), "Intercept"),
  prior(student_t(3,0,20), "sd"),
  prior(student_t(3,0,20), "sigma")
)

```

Fit the model 

```{r, eval = FALSE}


# run Bayesian model including phylogenetic tree
mod <- brm((logRMR)  ~ logM + T + therm # fixed coefficients
								+ (1 | gr(spp, cov = A)) + (1| spp_name), # random factors
  								family = gaussian(), 
  								data2 = list(A = A),
  								prior = priors,
  								sample_prior = TRUE, 
  								chains = 2, 
  								cores = 8, # cores to be used 8 in mac pro
  								iter = 3000, 
  								warmup = 500,
  								control = list(adapt_delta = 0.99, max_treedepth = 15), # Improve convergence
  								save_pars = save_pars(all = TRUE),
  								data = rmr_data)
  								

```


Check model convergence.

```{r, eval = FALSE}
# check model convergence 
plot(mod)
pp_check(mod)

# extract posteriors
posterior_samples <- posterior_samples(mod)

# get coefficients
summary(mod) 

# get bayesian p-value
p_map(mod)

```

Plot the fitted model 

```{r}

```




