---
title: "The relationship between critical ambient water temperature, body mass and thermal strategy"
author: "Andrew L Jackson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The relationship between critical ambient water temperature, body mass and thermal strategy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{tidyverse}
  \usepackage[utf8]{inputenc}
---


## Setup

```{r}
# load the predictFishRMR package 
library(predictFishRMR)

# for plotting and tibble manipulation
library(tidyverse)


```

## Sensitivity of our estimate on K

In this example we explore how our estimate of critical ambient temperature, above which we predict a fish to be unable to maintain long term thermal equilibrium. We generate estimates over a range of body masses for both thermal strategies of mesothermy and ectothermy. We use the mean K as predicted according to Nakamura's approach alongside a sensitivity whereby this value is assumed to vary by a factor of 2 in both directions (i.e. K is taken as both double and half its predicted value), thereby providing a 4-fold range. 

```{r}

m_range <- seq(100, 10000, length.out = 100)

Ta_crit_meso <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 1)
Ta_crit_ecto <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 1)

Ta_crit_meso_high <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 2)
Ta_crit_ecto_high <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 2)

Ta_crit_meso_low <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 0.5)
Ta_crit_ecto_low <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 0.5)

dd_crit <- tibble(mass = rep(m_range, 2), 
                      Ta_crit = c(Ta_crit_meso,   Ta_crit_ecto),
                      low = c(Ta_crit_meso_low,   Ta_crit_ecto_low), 
                      high = c(Ta_crit_meso_high, Ta_crit_ecto_high),
                      'Thermal Strategy' = c(rep("mesotherm", length(m_range)), 
                               rep("ectotherm", length(m_range))))



gg_crit <- ggplot(data = dd_crit, 
                  mapping = aes(x = mass, y = Ta_crit, 
                                color = `Thermal Strategy`)) + 
  geom_path() + 
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8) + 
  geom_ribbon(aes(ymin = low, 
                  ymax = high), 
              alpha = 0.1, 
              linetype = 2) + 
  ylab("Critical water temperature (C)") + 
  xlab("Mass (kg)")

print(gg_crit)

```

**Figure X.** The predicted maximum water temperature in which a fish can maintain heat balance as a function of body mass and physiological type. Note that these are analytical approximations.



