---
title: "Fit the phylogenetic regression of RMR by body mass"
author: "Andrew L Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit the phylogenetic regression of RMR by body mass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{rstan}
  %\VignetteDepends{brms}
  %\VignetteDepends{ape}
  %\VignetteDepends{parallel}
  \usepackage[utf8]{inputenc}
---


## Setup

```{r}
library(predictFishRMR)
library(rstan)
library(brms)
library(ape)
library(parallel) # run models in parallel

```


## Introduction

In this file we illustrate how the phylogenetic regression model of RMR against body mass for the two thermal strategies was fitted. Because the fitting process is computationally expensive and can take some time, the early part of this vignette has code chunks set to `eval = FALSE`. These were set to `eval = TRUE` and the resultant model fit was saved to an `.rda` file. This `.rda` object is loaded in place of re-running the models, but users can take a copy of the `*.Rmd` associated with the vignette and change the eval flags to re-run the analysis themselves. We also include here the random seed value used to generate the model we present below, although it is possible that the use of the `parallel` package might result in different values generate if the number of cores differs by machine, or if the parallel instances do not inherit the same random seed.

```{r}
# set the random seed.
set.seed(101)
```


## Import and clean data

The raw species and tree datasets have to be tidied somewhat to ensure data types are correct and species character strings match exactly. 


### Resting metabolic dataset 

Load the raw RMR dataset.  

```{r}

# locate the full directory on this local machine for the MandT data
# in this package
rmr_file_string <- system.file("extdata/MandT.csv",
                                  package = "predictFishRMR")

# load the data
rmr_data <- read.csv( rmr_file_string )
```

The RMR data needs some cleaning so that the species names agree with those in the tree file.

```{r}
rmr_data$spp <- gsub("_", " ", rmr_data$spp) #remove underscore

# some corrections are needed
rmr_data$spp[rmr_data$spp == "Scomberesocidae sp."] <- 
  "Scomberesocidae" # correct this species
rmr_data$spp[rmr_data$spp == "Ambassis sp"] <- 
  "Ambassis sp." # correct this species
rmr_data$spp[rmr_data$spp == "Hypoatherina sp"] <- 
  "Hypoatherina sp." # correct this species
rmr_data$spp[rmr_data$spp == "Monacanthidae sp"] <- 
  "Monacanthidae" # correct this species
rmr_data$spp[rmr_data$spp == "Monacanthidae sp."] <- 
  "Monacanthidae" # correct this species
rmr_data$spp[rmr_data$spp == "Carcharodon carcharius"] <- 
  "Carcharodon carcharias" # correct this species
rmr_data$spp[rmr_data$spp == "Stegostoma tigrinum"] <- 
  "Stegostoma fasciatum"
rmr_data <- subset(rmr_data, spp != "Caulophrynidae sp.") # this family is not included in the phylo tree
rmr_data <- rmr_data[order(rmr_data$spp),]
```

### Species names dataset

Load the species names data.

```{r}
# locate the full directory on this local machine for the species names data
# in this package
sp_file_string <- system.file("extdata/sp_list.csv",
                                  package = "predictFishRMR")

# load the file
sp_names <- read.csv(sp_file_string, sep = ";")
```

Process the species names data. 

```{r}
sp_names$spp <- gsub("_", " ", sp_names$spp) #remove underscore
sp_names$original.name <- gsub("_", " ", sp_names$original.name) #remove underscore
sp_names$sp_all <- gsub("_", " ", sp_names$sp_all) #remove underscore
sp_names$original.name[sp_names$original.name == "Carcharodon carcharius"] <- 
  "Carcharodon carcharias" # correct this species
sp_names$original.name[sp_names$original.name == "Stegostoma tigrinum"] <- 
  "Stegostoma fasciatum" # correct this species
```

### Phylogenetic tree dataset

```{r}
# locate the full directory on this local machine for the tree data
# in this package
tree_file_string <- system.file("extdata/shark_fish.trees",
                                  package = "predictFishRMR")

# load the file
tree <- read.tree(tree_file_string)
```

Process the tree tips for matching. 

```{r}
tree$tip.label <- gsub("_", " ", tree$tip.label) #remove underscore
tree$tip.label <- gsub("-", ",", tree$tip.label) #remove underscore
tree$tip.label[tree$tip.label == 
                 "Pseudocaranx dentex ,formaly Longirostrum delicatissimus"] <- 
  "Pseudocaranx dentex (formaly Longirostrum delicatissimus) "
tree$tip.label[tree$tip.label == "Sardinops sagax ,formaly S. caerulea"] <- 
  "Sardinops sagax (formaly S. caerulea)"
```

### Name matching and variable type classification

Check that the names match in the trait data and the species tree data.

```{r}
# Info for Stegostoma_tigrinum comes from Stegostoma_fasciatum
sp_tree <- unique(tree$tip.label)
sp <- unique(rmr_data$spp)

# check point: diff must be 0
setdiff(rmr_data$spp, tree$tip.label)
setdiff(rmr_data$spp, sp_names$original.name)

# convert some vectors to factor
rmr_data$method <- factor(rmr_data$method)
rmr_data$ram    <- factor(rmr_data$ram)
rmr_data$therm  <- factor(rmr_data$therm)

# ddrop incomplete rows
rmr_data        <- rmr_data[complete.cases(rmr_data),]
```

## Bayesian Linear Mixed / Phylogenetic Model

This chunk is not evaluated by default. 

```{r, eval = FALSE}
# construct covariance matrix of species (Hadfield & Nakagawa, 2010).
A <- ape::vcv.phylo(compute.brlen(tree,method="Grafen"))

# adding species as a factor into our dataset
rmr_data$spp_name<-factor(rmr_data$spp)

# Many species include serveral masurements. Thus, we use a bayesian models.
# General STAN specifications 
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

		
		
		# MODEL
		# We use a bayesian approach to study the relationship
		# across species (phylogeny) while taking into considration
		# intraspaecific variation and multiple observations for
		# several species.



# set model priors
priors =c(
  prior(normal(0,10), "b"),
  prior(normal(0,50), "Intercept"),
  prior(student_t(3,0,20), "sd"),
  prior(student_t(3,0,20), "sigma")
)

```

Fit the model. This chunk is not evaluated by default and instead the subsequent chunk will load the data. 

```{r, eval = FALSE}


# run Bayesian model including phylogenetic tree
mod <- brm((logRMR)  ~ logM + T + therm # fixed coefficients
								+ (1 | gr(spp, cov = A)) + (1| spp_name), # random factors
  								family = gaussian(), 
  								data2 = list(A = A),
  								prior = priors,
  								sample_prior = TRUE, 
  								chains = 3, 
  								cores = 8, # cores to be used 8 in mac pro
  								iter = 3000, 
  								warmup = 500,
  								control = list(adapt_delta = 0.99, max_treedepth = 15), # Improve convergence
  								save_pars = save_pars(all = TRUE),
  								data = rmr_data)
  								

```

Load the pre-run model data object and assign it the name "mod".

```{r}
data("fittedModelRMR")

mod <- fittedModelRMR


```



Plot histograms of the parameter posterior estimates alongside traceplots of the chains. 

```{r, eval = TRUE}
# Plot of histograms and traceplots
plot(mod)

```

Posterior predictive check plot

```{r}
pp_check(mod)
```



Calculate summaries of the posterior estimates including convergence checks. 

```{r}

# extract posteriors
posterior_samples <- as_draws(mod)

# get coefficients
summary(mod) 

# get bayesian p-value
# p_map(mod)
```


Save the fitted model. This chunk is not evaluated by default and the internal commands are commmented out. Users can uncomment and set to `eval = TRUE` if they wish to run this and save the object locally.

```{r, eval = FALSE}
save(fittedModelRMR, file = "fittedModelRMR.rda", 
     compress = "xz")
```




