---
title: "Fit the phylogenetic regression of RMR by body mass"
author: "Andrew L Jackson"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fit the phylogenetic regression of RMR by body mass}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{rstan}
  %\VignetteDepends{brms}
  %\VignetteDepends{ape}
  %\VignetteDepends{parallel}
  \usepackage[utf8]{inputenc}
---


## Setup

```{r}
library(predictFishRMR)
library(rstan)
library(brms)
library(ape)
library(parallel) # run models in parallel

```


## Introduction

In this file we illustrate how the phylogenetic regression model of RMR against body mass for the two thermal strategies was fitted. Because the fitting process is computationally expensive and can take some time, the early part of this vignette has code chunks set to `eval = FALSE`. These were set to `eval = TRUE` and the resultant model fit was saved to an `.rda` file. This `.rda` object is loaded in place of re-running the models, but users can take a copy of the `*.Rmd` associated with the vignette and change the eval flags to re-run the analysis themselves. We also include here the random seed value used to generate the model we present below, although it is possible that the use of the `parallel` package might result in different values generate if the number of cores differs by machine, or if the parallel instances do not inherit the same random seed.

```{r}
# set the random seed.
set.seed(101)
```


## Import and clean data

The raw species and tree datasets have to be tidied somewhat to ensure data types are correct and species character strings match exactly. 


### Resting metabolic dataset 

Load the raw RMR dataset.  

```{r}

# locate the full directory on this local machine for the MandT data
# in this package
rmr_file_string <- system.file("extdata/MandT.csv",
                                  package = "predictFishRMR")

# load the data
rmr_data <- read.csv( rmr_file_string )
```

The RMR data needs some cleaning so that the species names agree with those in the tree file.

```{r}

```

### Species names dataset

Load the species names data.

```{r}

```

Process the species names data. 

```{r}

```

### Phylogenetic tree dataset

```{r}
# locate the full directory on this local machine for the tree data
# in this package
tree_file_string <- system.file("extdata/shark_fish.trees",
                                  package = "predictFishRMR")

# load the file
tree <- read.tree(tree_file_string)
```

Process the tree tips for matching. 

```{r}

```

### Name matching and variable type classification

Check that the names match in the trait data and the species tree data.

```{r}

```

## Bayesian Linear Mixed / Phylogenetic Model

This chunk is not evaluated by default. 

```{r, eval = FALSE}

```

Fit the model. This chunk is not evaluated by default and instead the subsequent chunk will load the data. 

```{r, eval = FALSE}

  								

```

Load the pre-run model data object and assign it the name "mod".

```{r}
data("fittedModelRMR")

mod <- fittedModelRMR


```



Plot histograms of the parameter posterior estimates alongside traceplots of the chains. 

```{r, eval = TRUE}
# Plot of histograms and traceplots
plot(mod)

```

Posterior predictive check plot

```{r}
pp_check(mod)
```



Calculate summaries of the posterior estimates including convergence checks. 

```{r}

# extract posteriors
posterior_samples <- as_draws(mod)

# get coefficients
summary(mod) 

# get bayesian p-value
# p_map(mod)
```


Save the fitted model. This chunk is not evaluated by default and the internal commands are commmented out. Users can uncomment and set to `eval = TRUE` if they wish to run this and save the object locally.

```{r, eval = FALSE}
save(fittedModelRMR, file = "fittedModelRMR.rda", 
     compress = "xz")
```




