---
title: "The relationship between critical ambient water temperature, body mass and thermal strategy"
author: "Andrew L Jackson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The relationship between critical ambient water temperature, body mass and thermal strategy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{tidyverse}
  %\VignetteDepends{brms}
  \usepackage[utf8]{inputenc}
---


## Setup

```{r}
# load the predictFishRMR package 
library(predictFishRMR)

# for plotting and tibble manipulation
library(tidyverse)

# for extracting the coefficients of the brms regression model
library(brms)

```

## Sensitivity of our estimate on K

In this example we explore how our estimate of critical ambient temperature, above which we predict a fish to be unable to maintain long term thermal equilibrium. We generate estimates over a range of body masses for both thermal strategies of mesothermy and ectothermy. We use the mean K as predicted according to Nakamura's approach alongside a sensitivity whereby this value is assumed to vary by a factor of 2 in both directions (i.e. K is taken as both double and half its predicted value), thereby providing a 4-fold range. 

```{r}

# generate a sequence of mass for the x-axis of the plots
m_range <- seq(100, 10000, length.out = 100)

# load and retrieve the parameters for the RMR regression
# order of the parameters is:
# 1 = Intercept (gamma)
# 2 = alpha the coefficient of log10(mass)
# 3 = beta the coefficient of Body temperature
# 4 = the effect of mesothermy (psi)
data("fittedModelRMR")
pars <- brms::fixef(fittedModelRMR)[,1]

# calculate the mean predicted Ta_crit for the mass vector
Ta_crit_meso <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 1)
Ta_crit_ecto <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 1)

# calculate Ta_crit using 2*K in the calculations
Ta_crit_meso_high <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 2)
Ta_crit_ecto_high <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 2)

# calculate Ta_crit using K/2 in the calculations
Ta_crit_meso_low <- maxAmbientTemp(m_range, meso = TRUE, pars, preK = 0.5)
Ta_crit_ecto_low <- maxAmbientTemp(m_range, meso = FALSE, pars, preK = 0.5)

# assemble the data.frame (tibble) for plotting
dd_crit <- tibble(mass = rep(m_range, 2), 
                      Ta_crit = c(Ta_crit_meso,   Ta_crit_ecto),
                      low = c(Ta_crit_meso_low,   Ta_crit_ecto_low), 
                      high = c(Ta_crit_meso_high, Ta_crit_ecto_high),
                      'Thermal Strategy' = c(rep("mesotherm", length(m_range)), 
                               rep("ectotherm", length(m_range))))


# plot the curves including the sensitivity as a ribbon plot
gg_crit <- ggplot(data = dd_crit, 
                  mapping = aes(x = mass, y = Ta_crit, 
                                color = `Thermal Strategy`)) + 
  geom_path() + 
  scale_color_viridis_d(option = "D", begin = 0.2, end = 0.8) + 
  geom_ribbon(aes(ymin = low, 
                  ymax = high), 
              alpha = 0.1, 
              linetype = 2, 
              show.legend = FALSE) + 
  ylab("Critical water temperature (C)") + 
  xlab("Mass (kg)") + 
  ylim(c(0, 60)) + 
  xlim(c(0, 2500))

# print to screen
print(gg_crit)

```

**Figure X.** The predicted maximum water temperature in which a fish can maintain heat balance as a function of body mass and physiological type. Note that these are analytical approximations. The dotted lines represent a sensitivity analysis whereby the value of K used in the estimates was both doubled and halved.



